Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-19 WSSSPE2_report/WSSSPE2_review.tex
edits in 4 and 11

\item \textbf{Marlon Pierce: Patching It Up, Pulling It Forward.}
Pierce discussed how open open source is. Open software needs a diverse, openly
governed community behind it, just as it needs open licensing and a public code
repository. To probe the level of governance within open source projects, he and
his co-authors (Suresh Marru and Chris Mattman) suggested a contest to encourage
individual developers to submit patches and pull request to projects that are
important to them. This simple mechanism will expose several governance
mechanisms, such as how easy it is for independent developers to communicate
with project leadership, how projects accept and properly license third party
contributions, and how projects make decisions such as granting source tree
write access.

@@

\item \textbf{Abani Patra: Execute it.}
%Univ at Buffalo, SUNY
The primary focus of this presentation was to reinforce the value of
an available and easily accessible platform for executing scientific software
e.g. using platforms like Hubzero to access XSEDE or other computing resources.
Availability of such a platform for executing benchmark problems (even at small
scale) allows the developer community access to a reference implementation and
an easy way to train the larger user community. A second focus of this
presentation was the idea that much more attention needs to be paid to the
integrated use of simulation tools inside complex workflows. True usability
needs supporting the user community workflow. \todo{revise}

@@

\begin{table*}[t]
\centering
\caption{Top tweets tagged \#WSSSPE on Nov 16, 2014.}\label{tab:tweets}
  \begin{scriptsize}
  \begin{tabular}{l|l|r|r}
 \hline
    Author  &   Tweet  & Retweets &  Favorites
\\ \hline 
% Software Carpentry & Nov 17 & You call it ``project planning'' if it hasn't started yet,          & 8 & 1
%\\                          & &  and``software sustainability'' if it has and wasn't planned.     &    & 
%
% Neil P Chue Hong   &  Here's @SoftwareSaved guidance on Writing and using a software & 7 & 1
%\\     &  management plan used by EPSRC software grants    &    &  
% \\     &  http://w%ww.software.ac.uk/resources/guides/software-management-plans  &    & 
%
 Neil P Chue Hong  &  @jameshowison as well as software plans   & 4 & 7
\\   &   http://www.software.ac.uk/resources/guides/software-management-plans &    & 
\\   &   we provide a software evaluation tool:   &    & 
\\   &   http://www.software.ac.uk/online-sustainability-evaluation  &    & 
%
\\ Dr Tom Crick  & $56\%$ of UK researchers develop their own software $\rightarrow  140,000$   &  14 & 8
\\ & UK researchers write research software w$/$out any formal training &    & 
%
\\ Karthik Ram  &  ``Institutionalize metadata before metadata institutionalizes you'' & 8 & 6
%
\\ Josh Greenberg  &  ``1. retract any paper with bitrotten dependencies'' *mic drop* & 13 & 8
\\   &   ``2. add anyone who fixes bitrot as an author'' *mic drop*  &    & 
%
\\ Ethan White &  ``@rOpenSci is all about community... our measures of success  & 9 & 3
\\ &   [include] how many faces are up on our community page''   &    & 
%
\\ Ethan White & Daniel Katz talking about implementing transitive credit for  & 9 & 7
\\ &  software http://arxiv.org/abs/1407.5117  Work with @arfon  &    & 
% 
%\\ Aleksandra Pawlik  &  ``Tell us how you test your scientific software'' @gvwilson @swcarpentry  & 7 & 2
% 
\\ Aleksandra Pawlik  & Lack of training as one of the main barriers for sustainable software & 10 & 4
\\ &    at @Supercomputing. @swcarpentry @datacarpentry can fix that!  & & 
%
\\ Kaitlin Thaney  & My slides from this morning's keynote at & 11 & 12
\\ &  WSSSPE on Designing for Truth, Scale $+$ Sustainability:  & & 
\\ &   http://www.slideshare.net/kaythaney/     & & 
\\ &  designing-for-truth-scale-and-sustainability-wssspe2-keynote   & & 
%
\\ Neil P Chue Hong & @kaythaney shout out for @swcarpentry @datacarpentry & 9 & 4
\\ &  @rOpenSci @stilettofiend around open training activities for sustainability  & & 
%
%\\ Richard Littauer &  Sweet! @swcarpentry has had 4000+ learners in the past year. & 6 & 
%
\\ Neil P Chue Hong & For those interested in Github - Figshare/Zenodo integration, & 5 & 12
\\ & but want SWORD/DSpace/Fedora/ePrints see:  & & 
\\ & http://blog.stuartlewis.com/2014/09/09/github-to-repository-deposit/  & & 
%
\\ Hilmar Lapp & Re: adopting the unix philosophy, consider signing the Small Tools in & 7 & 6
\\ & Bioinformatics Manifesto: https://github.com/pjotrp/bioinformatics   & 
%
\\Andre Luckow &  ``Traditions last not because they are excellent, & 12 & 3 
\\ & but because influential people are averse to change...''  C. Sunstein     & & 
% 
\\ Dr Tom Crick &  ``Can I Implement Your Algorithm?'':  A Model for  Reproducible & 9 & 8
\\   &  Research Software http://arxiv.org/abs/1407.5981  & & 
%
\\Mozilla Science Lab & At a loose end this Sunday? Care about reproducibility, software  $+$ & 10 & 5
\\ &  \#openscience? Follow the    \#WSSSPE hashtag for more, live from New Orleans.  &  &
%
\\Kaitlin Thaney & I'm in New Orleans at \#WSSSPE , speaking at 9:50 ET on  scientific software & 9 & 9
\\ &   $+$ sustainability. Tune in! Live stream: http://ustre.am/17ddh & & 
%
\\   Daniel S. Katz & \#WSSSPE Agenda (Sunday):  & 10 & 1
\\ & http://wssspe.researchcomputing.org.uk/wssspe2/agenda/   &  &
\\ & URL for live stream of keynotes \& lightning talks: http://ustre.am/17ddh   &  &
\\ \hline
    \end{tabular}
    \end{scriptsize}
\end{table*}

@@

WSSSPE has developed an engaging online mechanism via the online social network
Twitter (hashtag ``\#WSSSPE''). There were substantially more tweets (messages)
during WSSSPE2 than WSSSPE1 or WSSSPE1.1. Some of the main points and highlights
in the meeting reverberated in Twitter. Table~\ref{tab:tweets} summarizes the
top tweets happened on the day of workshop, Nov 16, 2014,

>>>>

\item \textbf{Marlon Pierce: Patching It Up, Pulling It Forward.}
Pierce discussed how open open source is. Open software needs a diverse, openly
governed community behind it, just as it needs open licensing and a public code
repository. To probe the level of governance within open source projects, he and
his co-authors (Suresh Marru and Chris Mattmann) suggested a contest to encourage
individual developers to submit patches and pull request to projects that are
important to them. This simple mechanism will expose several governance
mechanisms, such as how easy it is for independent developers to communicate
with project leadership, how projects accept and properly license third party
contributions, and how projects make decisions such as granting source tree
write access.

@@

\item \textbf{Abani Patra: Execute it.}
%Univ at Buffalo, SUNY
The primary purpose of this presentation was to reinforce the value of
an available and easily accessible platform for executing scientific software,
e.g., HUBzero, to access XSEDE or other computing resources.
Such a platform for executing benchmark problems (even at small
scale) allows the developer community access to a reference implementation and
provides an easy way to train the larger user community. A second focus of this
presentation was the idea that much more attention needs to be paid to the
integrated use of simulation tools inside complex workflows. True usability
needs supporting the user community workflow. \todo{revise}

@@

\begin{table*}[t]
\centering
\caption{Top tweets tagged \#WSSSPE on Nov 16, 2014.}\label{tab:tweets}
  \begin{scriptsize}
  \begin{tabular}{l|l|r|r}
 \hline
    Author  &   Tweet  & Retweets &  Favorites
\\ \hline 
% Software Carpentry & Nov 17 & You call it ``project planning'' if it hasn't started yet,          & 8 & 1
%\\                          & &  and``software sustainability'' if it has and wasn't planned.     &    & 
%
% Neil P Chue Hong   &  Here's @SoftwareSaved guidance on Writing and using a software & 7 & 1
%\\     &  management plan used by EPSRC software grants    &    &  
% \\     &  http://w%ww.software.ac.uk/resources/guides/software-management-plans  &    & 
%
 Neil P Chue Hong  &  @jameshowison as well as software plans   & 4 & 7
\\   &   http://www.software.ac.uk/resources/guides/software-management-plans &    & 
\\   &   we provide a software evaluation tool:   &    & 
\\   &   http://www.software.ac.uk/online-sustainability-evaluation  &    & 
%
\\ Tom Crick  & $56\%$ of UK researchers develop their own software $\rightarrow  140,000$   &  14 & 8
\\ & UK researchers write research software w$/$out any formal training &    & 
%
\\ Karthik Ram  &  ``Institutionalize metadata before metadata institutionalizes you'' & 8 & 6
%
\\ Josh Greenberg  &  ``1. retract any paper with bitrotten dependencies'' *mic drop* & 13 & 8
\\   &   ``2. add anyone who fixes bitrot as an author'' *mic drop*  &    & 
%
\\ Ethan White &  ``@rOpenSci is all about community... our measures of success  & 9 & 3
\\ &   [include] how many faces are up on our community page''   &    & 
%
\\ Ethan White & Daniel Katz talking about implementing transitive credit for  & 9 & 7
\\ &  software http://arxiv.org/abs/1407.5117  Work with @arfon  &    & 
% 
%\\ Aleksandra Pawlik  &  ``Tell us how you test your scientific software'' @gvwilson @swcarpentry  & 7 & 2
% 
\\ Aleksandra Pawlik  & Lack of training as one of the main barriers for sustainable software & 10 & 4
\\ &    at @Supercomputing. @swcarpentry @datacarpentry can fix that!  & & 
%
\\ Kaitlin Thaney  & My slides from this morning's keynote at & 11 & 12
\\ &  WSSSPE on Designing for Truth, Scale $+$ Sustainability:  & & 
\\ &   http://www.slideshare.net/kaythaney/     & & 
\\ &  designing-for-truth-scale-and-sustainability-wssspe2-keynote   & & 
%
\\ Neil P Chue Hong & @kaythaney shout out for @swcarpentry @datacarpentry & 9 & 4
\\ &  @rOpenSci @stilettofiend around open training activities for sustainability  & & 
%
%\\ Richard Littauer &  Sweet! @swcarpentry has had 4000+ learners in the past year. & 6 & 
%
\\ Neil P Chue Hong & For those interested in Github - Figshare/Zenodo integration, & 5 & 12
\\ & but want SWORD/DSpace/Fedora/ePrints see:  & & 
\\ & http://blog.stuartlewis.com/2014/09/09/github-to-repository-deposit/  & & 
%
\\ Hilmar Lapp & Re: adopting the unix philosophy, consider signing the Small Tools in & 7 & 6
\\ & Bioinformatics Manifesto: https://github.com/pjotrp/bioinformatics   & 
%
\\Andre Luckow &  ``Traditions last not because they are excellent, & 12 & 3 
\\ & but because influential people are averse to change...''  C. Sunstein     & & 
% 
\\ Tom Crick &  ``Can I Implement Your Algorithm?'':  A Model for  Reproducible & 9 & 8
\\   &  Research Software http://arxiv.org/abs/1407.5981  & & 
%
\\Mozilla Science Lab & At a loose end this Sunday? Care about reproducibility, software  $+$ & 10 & 5
\\ &  \#openscience? Follow the    \#WSSSPE hashtag for more, live from New Orleans.  &  &
%
\\Kaitlin Thaney & I'm in New Orleans at \#WSSSPE , speaking at 9:50 ET on  scientific software & 9 & 9
\\ &   $+$ sustainability. Tune in! Live stream: http://ustre.am/17ddh & & 
%
\\   Daniel S. Katz & \#WSSSPE Agenda (Sunday):  & 10 & 1
\\ & http://wssspe.researchcomputing.org.uk/wssspe2/agenda/   &  &
\\ & URL for live stream of keynotes \& lightning talks: http://ustre.am/17ddh   &  &
\\ \hline
    \end{tabular}
    \end{scriptsize}
\end{table*}

@@

WSSSPE actively used the online social network
Twitter, with hashtag ``\#WSSSPE''. There were substantially more tweets (messages)
during WSSSPE2 than WSSSPE1 or WSSSPE1.1. Some of the main points and highlights
in the meeting are shown in Table~\ref{tab:tweets}, which summarizes the
top \#WSSSPE tweets from the day of workshop.
